---
title: 'Lake Erie HABs Community Ecology Manuscript'
output:
  html_document:
    toc: true
---
    




```{r global_options, include=FALSE}

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.height = 5, 
  fig.align = 'center', 
  fig.path = 'Figs/',
  warning = FALSE, 
  message = FALSE
)

```


```{r load libraries}

# Load libraries
setwd("~/chabs/miseq_may2015/analysis")
library(phyloseq)
library(DESeq2)
library(ggplot2)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(gridExtra)
library(RColorBrewer)
library(vegan)
library(magrittr)
library(knitr)
library(cowplot)
library(gtable)
library(multcompView)
library(pgirmess)
library(pander)
source("habs_functions.R")
source("~/git_repos/MicrobeMiseq/R/miseqR.R")
theme_set(theme_bw())

stationColors <- c("#1f78b4", "#33a02c", "#fb9a99")
```

```{r mothur import, echo = FALSE}

# Import mothur files and sample metadata
sharedfile = "mothur/chabs.shared"
taxfile = "mothur/chabs-fwdb-silva.taxonomy"
mapfile = "subdata-metadata.csv"

mothurdata = import_mothur(mothur_shared_file = sharedfile, 
	mothur_constaxonomy_file = taxfile)

# Add the OTU number as a column in the taxonomy file
tax_table(mothurdata) <- cbind(tax_table(mothurdata), 
	row.names(tax_table(mothurdata)))

# Rename the taxonomy columns
colnames(tax_table(mothurdata)) <- 
  c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Rank7", "Rank8", "Species")
  
# Import sample metadata and transform into sample_data class
map <- read.csv(mapfile)

```


```{r}

# Convert map to sample_data class
# Merge mothurdata object with sample_data
map <- sample_data(map)
rownames(map) <- map$SampleID
mothur.merge <- merge_phyloseq(mothurdata, map)
  

# Filter out non-samples (i.e. water, mock) and samples from intensive cruises.
erie.subset <-
  mothur.merge %>%
    subset_taxa(
      Kingdom == "Bacteria" &
      Family != "mitochondria" &
      Class != "Chloroplast"
    ) %>%
    subset_samples(
      Type == "sample" & 
      !(Date %in% c("5/27", "6/10", "8/11", "11/3"))
    )

erie <- subset_samples(erie.subset, Fraction == "CNA")
    

# Also prune out taxa which were only present in removed samples
erie <- prune_taxa(taxa_sums(erie) > 0, erie)


```


```{r normalization}

# Scale reads to even depth 
erieScale <-
  erie %>%
    scale_reads(n = 15000, round = "round") 


n = 15000
thresh = 0.0001

# Prune low abundance taxa using thresh as mean relative abundance
tax_mean <- taxa_sums(erieScale)/nsamples(erieScale)
erieScale.0001 <- prune_taxa(tax_mean > thresh*n, erieScale)

```


```{r}
# Import metadata file with nutrients, pigments and toxin
nutrient <- read.csv("other/nutrient_cleaned.csv")


# Format
nutrient <-
  nutrient %>%
    filter(!(Date %in% c("5/27", "6/10", "11/3"))) %>%
    order_dates() 

```
              
              

## Figure 2: Bloom and Bacterial Composition 

```{r figure2, fig.height=12, fig.width=8.5}

# Calculate relative abundance of Cyanobacteria at each date
cyanoAbundance <- 
  erieScale %>%
  tax_glom(taxrank = "Phylum") %>%
  transform_sample_counts(function(x){x/sum(x)}) %>%
  subset_taxa(Phylum == "Cyanobacteria") %>%
  psmelt() %>%
  rename(Cyanobacteria = Abundance) %>%
  select(Cyanobacteria, Date, Station)

# Merge with Nutrient df
nutrientCyano <-
  nutrient %>%
    left_join(cyanoAbundance, by = c("Station", "Date")) %>%
    order_dates()

# Create a new df which changes Phycocyanin levels above 50 to 50 
# so that boxplots are more readable 
nutrientCyanoAdj <-
  nutrientCyano %>%
  mutate(Phycocyanin = ifelse(Phycocyanin < 50, Phycocyanin, 50))




#################### Lineplots ####################

# Function to create a lineplot given a df and column name
make_bloomplot <- function(df, yvar, ylab) {
  ggplot(df, aes_string(x = "Date", y = yvar, group = "Station", linetype = "Station")) +
    geom_line(size = .8) +
    ylab(ylab) +
    scale_x_discrete(
      breaks = c("7/8", "8/4", "9/2", "10/6"),
      labels = c("Jul", "Aug", "Sep", "Oct"),
      drop = FALSE
    ) +
    theme(
      axis.title.y = element_text(size = 10),
      axis.title.x = element_blank(),
      panel.grid.major.x = element_blank(),
      plot.title = element_text(size = 13, vjust = 1.5),
      plot.margin = unit(c(.25, 1, .25, .5), "cm"),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 9)
    )
}


bloomVars <- c("Phycocyanin", "Chla", "ParMC", "CDOM", "Cyanobacteria")
bloomPlots <- list()

# Loop through stations and bloomvars and make a ggplot lineplot. 
for (station in levels(nutrientCyano$Station)) {
  nutrientCyanoStation <- filter(nutrientCyano, Station == station)
  for (var in bloomVars) {
    if (var == "Cyanobacteria") {
      thisylab <- "Rel. Abundance"
    } else if (var == "CDOM") {
      thisylab <- "a305 (m^-1)"
    } else {
      thisylab <- "ug/L"
    }
    plot <- make_bloomplot(nutrientCyanoStation, var, thisylab) + ggtitle(var)
    plot <- plot + theme(legend.position = "none")
    bloomPlots[[station]][[var]] <- plot
  }
}


################## Boxplots #####################

bloomBoxplots <- list()

# Function to make a boxplot of all stations
# Takes a df, column name, and chosen y-axis label as inputs
makebloom_boxplots <- function(df, var, ylab, labelsdf, labelylab) {
  ggplot(df, aes_string(x = "Station", y = var)) + 
    geom_boxplot() + 
    ggtitle(var) + 
    ylab(ylab) +
    scale_x_discrete(
      labels = c("N1", "N2", "O")
    ) +
    theme(
      axis.text.x = element_text(size = 9),
      axis.title.x = element_blank(),
      axis.title.y = element_text(size = 10),
      legend.position = "none",
      plot.title = element_text(size = 13, vjust = 1.5), 
      plot.margin = unit(c(.5, .5, .25, 1), "cm")
    )
}



# Loop through bloomvars and create a boxplot for each comparing all stations
for (var in bloomVars) {
  if (var == "Cyanobacteria") {
      thisylab <- "Rel. Abundance"
    } else if (var == "CDOM") {
      thisylab <- "a305 (m^-1)"
    } else {
      thisylab <- "ug/L"
    }
  plot <- makebloom_boxplots(df = nutrientCyanoAdj, var = var, ylab = thisylab, labelsdf = boxplotLetters[[var]], labelylab = maxY[[var]])
 bloomBoxplots[[var]] <- plot
  
}


####################### Compiled Plot ##########################

# Grab the grobs for each station since only n1 is used as a main figure
# n2 and offshore will be in supplement
bloomgrobsN1 <- llply(bloomPlots$nearshore1, ggplotGrob)
bloomgrobsN2 <- llply(bloomPlots$nearshore2, ggplotGrob)
bloomgrobsO <- llply(bloomPlots$offshore, ggplotGrob)

# Grab the grobs for the boxplots
boxplotGrobs <- llply(bloomBoxplots, ggplotGrob)

# Plot A: lineplots
plota <- do.call("rbind", c(bloomgrobsN1 , size = "last"))

# Plot B: boxplots
plotb <- do.call("rbind", c(boxplotGrobs, size = "last"))


ggdraw() +
  draw_plot(cbind(plota, plotb, size = "last"), 
            x = 0.01, y = 0.01, width = .98, height = .98) +
  draw_plot_label(c("A", "B"), c(0.02, .5), c(1, 1), size = 20) 
  
```

# Figure 2 Statistics
```{r figure2stats}

# Calculate correlation between Chla and phycocyanin for all sites
cor.test(
  x = nutrient$Chla, 
  y = nutrient$Phycocyanin, 
  alternative = "greater", 
  method = "pearson"
)


# Calculate Microcystis relative abundances
microcystisAbundance <- 
  erieScale %>%
  transform_sample_counts(function(x){x/sum(x)}) %>%
  subset_taxa(Genus == "Microcystis") %>%
  tax_glom(taxrank = "Genus") %>%
  psmelt() %>%
  rename(Microcystis = Abundance) %>%
  select(Microcystis, Station, Date)
  

nutrientCyano <- 
  nutrientCyano %>%
    left_join(microcystisAbundance, by = c("Station", "Date"))

# Kruskal-wallis and posthoc test

kruskalResults <- list()

for (var in c(bloomVars, "Microcystis")) {
  kruskalResults[[var]] <- kruskal.test(
    formula = reformulate("Shore", var), 
    data = nutrientCyano)$p.value
}

kruskalFdr <- p.adjust(unlist(kruskalResults), method = "fdr")

kruskalFdr

kruskalPostResults <- list()
for (var in c(bloomVars, "Microcystis")) {
  kruskalPostResults[[var]] <- kruskalmc(
    resp = nutrientCyano[ ,var], 
    categ = nutrientCyano$Station)
}


# boxplotLettersBoolean <- 
#   lapply(
#     kruskalPostResults, function(x) {x$dif.com$difference}
#   )
# 
# stationPairwise = c("nearshore1-nearshore2", "nearshore1-offshore", "nearshore2-offshore")
# 
# names(boxplotLettersBoolean$Phycocyanin) <- stationPairwise
# names(boxplotLettersBoolean$Chla) <- stationPairwise
# names(boxplotLettersBoolean$ParMC) <- stationPairwise
# names(boxplotLettersBoolean$CDOM) <- stationPairwise
# names(boxplotLettersBoolean$Cyanobacteria) <- stationPairwise
# 
# boxplotLetters <- list()
# 
# boxplotLetters[["Phycocyanin"]] <- data.frame(labels = multcompLetters(boxplotLettersBoolean[["Phycocyanin"]], Letters = c(letters, LETTERS, "."))$Letters, station = Stations)
# boxplotLetters[["Chla"]] <- data.frame(labels = multcompLetters(boxplotLettersBoolean[["Chla"]], Letters = c(letters, LETTERS, "."))$Letters, station = Stations)
# boxplotLetters[["ParMC"]] <- data.frame(labels = multcompLetters(boxplotLettersBoolean[["ParMC"]], Letters = c(letters, LETTERS, "."))$Letters, station = Stations)
# boxplotLetters[["CDOM"]] <- data.frame(labels = multcompLetters(boxplotLettersBoolean[["CDOM"]], Letters = c(letters, LETTERS, "."))$Letters, station = Stations)
# boxplotLetters[["Cyanobacteria"]] <- data.frame(labels = multcompLetters(boxplotLettersBoolean[["Cyanobacteria"]], Letters = c(letters, LETTERS, "."))$Letters, station = Stations)
# 
# # max y positions
# maxY$Phycocyanin = 55
# maxY$Chla = 60
# maxY$ParMC = 28
# maxY$CDOM = 8
# maxY$Cyanobacteria = .32

```


<br>

## PERMANOVA

```{r}
# Remove dates for which we are missing samples for any of the sites
stationsDateSubset <- subset_samples(erieScale, Date != "9/29")

# Calculate bray-curtis distance
bdist <- phyloseq::distance(physeq = stationsDateSubset, method = "bray")

# Convert sample_data to df
sampledf <- data.frame(sample_data(stationsDateSubset))

# Run adonis on sampling date
adonis(bdist ~ Date, data = sampledf)
beta <- betadisper(bdist, sampledf$Date)
permutest(beta)

# Run adonis on shore + station 
adonis(bdist ~ Shore + Station, data = sampledf)
beta <- betadisper(bdist, sampledf$Shore)
permutest(beta)

```
 

## Figure 3: Cyanobacteria community and OTU dynamics

```{r figure3, fig.height=12, fig.width=12}

# Subset to cyanobacteria and scale internally
cyanos <- 
  erie %>%
  subset_taxa(Class == "Cyanobacteria") %>%
  scale_reads(round = "round")



############################## PC plots ##########################

# Ordinate
cyanosPcoa <- ordinate(
  physeq = cyanos, 
  method = "PCoA", 
  distance = "bray"
)

# Determine which PC's exceed broken stick model
which(cyanosPcoa$values$Rel_corr_eig > cyanosPcoa$values$Broken_stick)

# Create new data frame with sample metadata and significant PCs
cyanosPcDF <- data.frame(
  SampleID = rownames(cyanosPcoa$vectors), 
  PC1 = cyanosPcoa$vectors[ ,1], 
  PC2 = cyanosPcoa$vectors[ ,2], 
  row.names = NULL
)

# Create a dataframe with sample metadata and sample coordinates on PC1 and PC2 for cyanos
cyanoDF <- 
    data.frame(sample_data(erie)) %>%
    left_join(cyanosPcDF, by = "SampleID")  %>%
    order_dates() 

cyanoPC <- c("PC1", "PC2")
cyanoRel_corr_eigs <- signif(cyanosPcoa$values$Rel_corr_eig[1:2] *100, 2)
names(cyanoRel_corr_eigs) <- cyanoPC

cyanosPcPlots <- list()

make_pc_plot <- function(df, pc, eigs) {
  ggplot(df, aes_string(x = "Date", y = pc, group = "Station", color = "Station")) +
      geom_line(size = 1.2) + 
      scale_color_manual(values = stationColors) +
      scale_x_discrete(
        breaks = c("7/8", "8/4", "9/2", "10/6"),
        labels = c("Jul", "Aug", "Sep", "Oct"),
        drop = FALSE
      ) +  
      ylab(paste(pc, " ", eigs[pc], "%")) +
      theme(
        axis.title.x = element_blank(),
        axis.text.x = element_text(
          angle = 45, hjust = 1, vjust = 1
        ),
        legend.position = "none",
        plot.margin = unit(c(0, 0.2, 0, 0.2), "cm"),
        axis.title.y = element_text(size = 13, face = "bold")
      ) 
}



# For loop to create PC plots
for (pc in cyanoPC) {
  cyanosPcPlots[[pc]] <- make_pc_plot(cyanoDF, pc, cyanoRel_corr_eigs)
}


####################### OTU plots #########################

make_otu_plot <- function(df, otu, taxrank) {
  ggplot(df, aes(x = Date, y = Abundance, group = Station, color = Station)) +
    geom_line() +
    ggtitle(paste(df[1, taxrank], otu)) +
    ylab("rel. abund") + 
    scale_color_manual(values = stationColors) +
    scale_x_discrete(
      breaks = c("7/8", "8/4", "9/2", "10/6"),
      labels = c("Jul", "Aug", "Sep", "Oct"),
      drop = FALSE
    ) +
    xlab("") + 
    theme(
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 9),
      axis.title.y = element_text(size = 8),
      legend.position = "none",
      plot.margin = unit(c(0, 0.3, 0, 0.2), "cm"),
      plot.title = element_text(size = 10, face = "bold")
    ) 
}

## Select only cyano OTUs that have mean relative abundace > 0.0001
n = 15000
thresh = 0.0001

# Prune low abundance taxa using thresh as mean relative abundance
tax_mean <- taxa_sums(erieScale)/nsamples(erieScale)
eriePrune.0001 <- prune_taxa(tax_mean > thresh*n, erieScale)

cyanoOtus <-
  eriePrune.0001 %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  subset_taxa(Class == "Cyanobacteria") %>%
  psmelt() %>%
  order_dates()

cyanoOtuNames <- levels(cyanoOtus$Species)

cyanoOtuPlots <- list()

for (otu in cyanoOtuNames) {
  df.otu <- filter(cyanoOtus, OTU == otu)
  cyanoOtuPlots[[otu]] <- make_otu_plot(df = df.otu, otu = otu, taxrank = "Genus")
}

##################### Compile plots ##########################

# grab grobs for PC plots
cyanoPcGrobs <- llply(cyanosPcPlots, ggplotGrob)

# Plot A: PC plots
plota <- do.call("cbind", c(cyanoPcGrobs, size = "first"))

# Plot B: Cyano OTUs
plotb <- do.call("arrangeGrob", c(cyanoOtuPlots, ncol = 3))

ggdraw() +
  draw_plot(plota,
            x = 0.03, y = 0.7, width = .92, height = .26) +
  draw_plot(plotb, 
            x = 0.03, y = 0.01, width = .92, height = 0.66) +
  draw_plot_label(c("A", "B"), c(0.0, 0.0), c(1, 0.72), size = 20) 




```

## Table 1: Correlations between station profiles on PC axes
```{r}

Stations = c("nearshore1", "nearshore2", "offshore")

make_station_braypc <- function(braypcDF, station) {
  stationPC <-
    braypcDF %>%
      filter(Date != "9/29") %>%
      filter(Station == station)
  
  return(stationPC)
}

# Function to run correlation tests
test_pc_corrs <- function(pcDF, pc, method) {
  
  n1Vn2 <- cor.test(pcDF[["nearshore1"]][[pc]], pcDF[["nearshore2"]][[pc]], method = method)
  n1Vo <- cor.test(pcDF[["nearshore1"]][[pc]], pcDF[["offshore"]][[pc]], method = method)
  n2Vo <- cor.test(pcDF[["nearshore2"]][[pc]], pcDF[["offshore"]][[pc]], method = method)
  
  d <- data.frame(r = c(n1Vn2$estimate, n2Vo$estimate, n1Vo$estimate), 
                  p = c(n1Vn2$p.value, n2Vo$p.value, n1Vo$p.value),
                  row.names = c("n1 v n2", "n1 v o", "n2 v o")
                  )
  return(d)
  
}


# Make cyano stations 
pcCyanoStation <- list()

for (station in Stations) {
  pcCyanoStation[[station]] <- make_station_braypc(cyanoDF, station)
}

cyanosSpearman <- list()
for (pc in c("PC1", "PC2")) {
  cyanosSpearman[[pc]] <- test_pc_corrs(
    pcDF = pcCyanoStation, 
    pc = pc, 
    method = "spearman"
  )
}

cyanosSpearmanCorrected <- ldply(
  cyanosSpearman, 
  function(x) x %>% mutate(p = p.adjust(p, method = "fdr"))
)

cyanosPearson <- list()

for (pc in c("PC1", "PC2")) {
  cyanosPearson[[pc]] <- test_pc_corrs(
    pcDF = pcCyanoStation, 
    pc = pc, 
    method = "pearson"
  )
}

cyanosPearsonCorrected <- ldply(
  cyanosPearson, 
  function(x) x %>% mutate(p = p.adjust(p, method = "fdr"))
)

cyanosPearsonCorrected$Comparison <- rep(c("N1 v N2", "N1 v O", "N2 v O"), 2)


emphasize.strong.cells(which(cyanosPearsonCorrected < 0.05, arr.ind = TRUE))
pander(cyanosPearsonCorrected)     

```


# Figure 4: Heterotroph community and OTU dynamics



```{r}

# Subset to Non-Cyanos and scale internally
nonCyano <- 
  erie %>%
  subset_taxa(Class != "Cyanobacteria") %>%
  scale_reads(round = "round")

# Ordinate
nonCyanoPcoa <- ordinate(
  physeq = nonCyano, 
  method = "PCoA", 
  distance = "bray"
)

# Determine which PC's exceed broken stick model
which(nonCyanoPcoa$values$Rel_corr_eig > nonCyanoPcoa$values$Broken_stick)

## Create dataframe with PC's that exceed model 
nonCyanoPC <- data.frame(
  SampleID = rownames(nonCyanoPcoa$vectors), 
  PC1 = nonCyanoPcoa$vectors[ ,1], 
  PC2 = nonCyanoPcoa$vectors[ ,2], 
  PC3 = -nonCyanoPcoa$vectors[ ,3], 
  row.names = NULL
)

# Join PC data with sample metadata
nonCyanoDF <- 
  data.frame(sample_data(erie)) %>%
    left_join(nonCyanoPC, by = "SampleID")  %>%
    order_dates() 


####################### Plot A #########################

nonCyanoPCs <- c("PC1", "PC2", "PC3")
nonCyanoRelCorrEigs <- signif(nonCyanoPcoa$values$Rel_corr_eig[1:3] *100, 2)
names(nonCyanoRelCorrEigs) <- nonCyanoPCs

nonCyanoPCPlots <- list()

# For loop to create PC plots
for (PC in nonCyanoPCs) {
  nonCyanoPCPlots[[PC]] <- make_pc_plot(nonCyanoDF, PC, eigs = nonCyanoRelCorrEigs)
}
```




```{r}
## Differentially abundant OTUs along PCs

# Function to run deseq2 differential abundance analysis. 
# Takes a phyloseq object and factor column in sample_data as arguments
# Returns a df with differentially abundant taxa
get_PC_deseq_OTUs <- function(physeq, var) {
  
 station.deseq <- phyloseq_to_deseq2(physeq, reformulate(var))
 station.deseq = DESeq(station.deseq, test = "Wald", fitType = "parametric")

  res = data.frame(results(station.deseq, cooksCutoff = FALSE))
  res = data.frame(OTU = row.names(res), res)
  alpha = 0.05
  sigtab <-
    res %>%
      filter(padj < alpha) 

  sigtab.station = cbind(sigtab, as(tax_table(physeq)[sigtab$OTU, ], "matrix"))
  
  return(sigtab.station)
}


# Create factor levels for samples on PCs 
nonCyanoDF <-
  nonCyanoDF %>%
    mutate(PC1group = ifelse(PC1 > 0, "2", "1")) %>%
    mutate(PC2group = ifelse(PC2 > 0, "2", "1")) %>%  
    mutate(PC3group = ifelse(PC3 > 0, "2", "1"))

# Keep OTUs above average relative abundance of 5 reads
a <- apply(otu_table(erie), MARGIN = 1, function(x) {mean(x)} )
keep <- a > 5
eriePrune <- prune_taxa(keep, erie)

nonCyano <- 
  eriePrune %>%
    subset_taxa(Phylum != "Cyanobacteria")

# Add PC groups into phyloseq sample_data.
sample_data(nonCyano)$PC1group <- nonCyanoDF$PC1group
sample_data(nonCyano)$PC2group <- nonCyanoDF$PC2group
sample_data(nonCyano)$PC3group <- nonCyanoDF$PC3group


# run deseq analyses for each station seperately for PC1, PC2, PC3


################ nearshore1 ###################

nonCyanoN1 <-
    nonCyano %>%
      subset_samples(Station == "nearshore1")

pc1DeseqN1 <- get_PC_deseq_OTUs(nonCyanoN1, var = "PC1group")
pc2DeseqN1 <- get_PC_deseq_OTUs(nonCyanoN1, var = "PC2group")
pc3DeseqN1 <- get_PC_deseq_OTUs(nonCyanoN1, var = "PC3group")

############################# nearshore2 ###############################

nonCyanoN2 <-
    nonCyano %>%
    subset_samples(Station == "nearshore2")

pc1DeseqN2 <- get_PC_deseq_OTUs(nonCyanoN2, var = "PC1group")
pc2DeseqN2 <- get_PC_deseq_OTUs(nonCyanoN2, var = "PC2group")
pc3DeseqN2 <- get_PC_deseq_OTUs(nonCyanoN2, var = "PC3group")


################# offshore ###############
nonCyanoO <-
  nonCyano %>%
    subset_samples(Station == "offshore")

pc1DeseqO <- get_PC_deseq_OTUs(nonCyanoO, var = "PC1group")
pc2DeseqO <- get_PC_deseq_OTUs(nonCyanoO, var = "PC2group")
pc3DeseqO <- get_PC_deseq_OTUs(nonCyanoO, var = "PC3group")


```


```{r}

############## Plot b #########################

# Filter to results that are positive along PC1
pc1DeseqN1Pos <- filter(pc1DeseqN1, log2FoldChange > 0)
pc1DeseqN2Pos <- filter(pc1DeseqN2, log2FoldChange > 0)
pc1DeseqOPos <- filter(pc1DeseqO, log2FoldChange > 0)

# Find intersection of results from all three sites
pc1IntersectPos <- intersect(
  intersect(pc1DeseqN1Pos$Species, pc1DeseqN2Pos$Species),
  pc1DeseqOPos$Species
)

nonCyanoOtuPlotsPos <- list()

for (otu in pc1IntersectPos) {
  otuMelt <-
  erieScale %>%
    transform_sample_counts(function(x){x/sum(x)}) %>%
    subset_taxa(Species == otu) %>%
    psmelt() %>%
    order_dates()
  
  nonCyanoOtuPlotsPos[[otu]] <- make_otu_plot(df = otuMelt, otu, taxrank = "Class")
}

pc1PosTax <- data.frame(tax_table(erie)[pc1IntersectPos, ])
pc1PosTax$type = rep("PC1-pos", nrow(pc1PosTax))
```

```{r}
############## Plot c #########################
# Filter to results that are positive along PC1
pc1DeseqN1Neg <- filter(pc1DeseqN1, log2FoldChange < 0)
pc1DeseqN2Neg <- filter(pc1DeseqN2, log2FoldChange < 0)
pc1DeseqONeg <- filter(pc1DeseqO, log2FoldChange < 0)

# Find intersection of results from all three sites
pc1NegIntersect <- intersect(pc1DeseqN1Neg$Species, pc1DeseqN2Neg$Species)
pc1NegIntersect <- intersect(pc1NegIntersect, pc1DeseqONeg$Species)

nonCyanoOtuPlotsNeg <- list()

for (otu in pc1NegIntersect) {
  otuMelt <-
  erieScale %>%
    transform_sample_counts(function(x){x/sum(x)}) %>%
    subset_taxa(Species == otu) %>%
    psmelt() %>%
    order_dates()
  
  nonCyanoOtuPlotsNeg[[otu]] <- make_otu_plot(df = otuMelt, otu, taxrank = "Class")
}

pc1NegTax <- data.frame(tax_table(erie)[pc1NegIntersect, ])

pc1NegTax$type = rep("PC1-neg", nrow(pc1NegTax))
```


```{r fig4, fig.height=12, fig.width=12}

## Compile plots

# grab grobs for PC plots
pcGrobs <- llply(nonCyanoPCPlots, ggplotGrob)

# Plot A: PC plots
plota <- do.call("cbind", c(pcGrobs, size = "last"))

# Plot B: OTU's that become overabundant during bloom
plotb <- do.call("arrangeGrob", c(nonCyanoOtuPlotsPos, ncol = 4))

# Plot C: OTU's that are reduced during bloom
plotc <- do.call("arrangeGrob", c(nonCyanoOtuPlotsNeg, ncol = 4))

ggdraw() +
  draw_plot(plota,
            x = 0.03, y = 0.72, width = 0.92, height = 0.25) +
  draw_plot(plotb, 
            x = 0.03, y = 0.15, width = 0.92, height = 0.54) +
  draw_plot(plotc,
            x = 0.03, y = 0.0, width = 0.92, height = 0.14) +
  draw_plot_label(c("A", "B", "C"), c(0, 0, 0), c(1, 0.72, 0.16), size = 20) 


```

```{r}

# Filter to results that are positive along PC2
pc2DeseqN1Pos <- filter(pc2DeseqN1, log2FoldChange > 0)
pc2DeseqN2Pos <- filter(pc2DeseqN2, log2FoldChange > 0)
pc2DeseqOPos <- filter(pc2DeseqO, log2FoldChange > 0)

# Find intersection of results from all three sites
pc2IntersectPos <- intersect(
  intersect(pc2DeseqN1Pos$Species, pc2DeseqN2Pos$Species),
  pc2DeseqOPos$Species
)

pc2PosTax <- data.frame(tax_table(erie)[pc2IntersectPos, ])
pc2PosTax$type = rep("PC2-pos", nrow(pc2PosTax))

# Filter to results that are positive along PC1
pc2DeseqN1Neg <- filter(pc2DeseqN1, log2FoldChange < 0)
pc2DeseqN2Neg <- filter(pc2DeseqN2, log2FoldChange < 0)
pc2DeseqONeg <- filter(pc2DeseqO, log2FoldChange < 0)

# Find intersection of results from all three sites
pc2IntersectNeg <- intersect(
  intersect(pc2DeseqN1Neg$Species, pc2DeseqN2Neg$Species),
  pc2DeseqONeg$Species
)

pc2NegTax <- data.frame(tax_table(erie)[pc2IntersectNeg, ])

pc2NegTax$type = rep("PC2-neg", nrow(pc2NegTax))

```

```{r}
# # Filter to results that are positive along PC3
# pc3DeseqN1Pos <- filter(pc3DeseqN1, log2FoldChange > 0)
# pc3DeseqN2Pos <- filter(pc3DeseqN2, log2FoldChange > 0)
# pc3DeseqOPos <- filter(pc3DeseqO, log2FoldChange > 0)
# 
# # Find intersection of results from all three sites
# pc3IntersectPos <- intersect(
#   intersect(pc3DeseqN1Pos$Species, pc3DeseqN2Pos$Species),
#   pc3DeseqOPos$Species
# )
# 
# pc3PosTax <- data.frame(tax_table(erie)[pc3IntersectPos, ])
# 
# pc3PosTax$type = rep("PC3-pos", nrow(pc3PosTax))
# 
# 
# # Filter to results that are positive along PC3
# pc3DeseqN1Neg <- filter(pc3DeseqN1, log2FoldChange < 0)
# pc3DeseqN2Neg <- filter(pc3DeseqN2, log2FoldChange < 0)
# pc3DeseqONeg <- filter(pc3DeseqO, log2FoldChange < 0)
# 
# # Find intersection of results from all three sites
# pc3IntersectNeg <- intersect(
#   intersect(pc3DeseqN1Neg$Species, pc3DeseqN2Neg$Species),
#   pc3DeseqONeg$Species
# )
# 
# pc3NegTax <- tax_table(erie)[pc3IntersectNeg, ]
# 
# pc3NegTax
# 
# 
# #deseqTax <- rbind(pc1PosTax, pc1NegTax, pc2PosTax, pc2NegTax, pc3PosTax)
#deseqTax$all = paste(deseqTax$Phylum,"; ", 
#                      deseqTax$Class,"; ",
#                      deseqTax$Order,"; ",
#                      deseqTax$Family,"; ",
#                      deseqTax$Genus,"; ",
#                      deseqTax$Rank7,"; ",
#                      deseqTax$Rank8,"; ",
#                      sep = "")

#write.csv(deseqTax, "deseq-taxa.csv")

```


```{r}

pcNonCyanoStation <- list()

for (station in Stations) {
  pcNonCyanoStation[[station]] <- make_station_braypc(nonCyanoDF, station)
}

nonCyanosSpearman <- list()
for (pc in nonCyanoPCs) {
  nonCyanosSpearman[[pc]] <- test_pc_corrs(
    pcDF = pcNonCyanoStation, 
    pc = pc, 
    method = "spearman"
  )
}

nonCyanosSpearmanCorrected <- ldply(nonCyanosSpearman, function(x) x %>% mutate(p = p.adjust(p, method = "fdr")))

nonCyanosPearson <- list()

for (pc in c("PC1", "PC2", "PC3")) {
  nonCyanosPearson[[pc]] <- test_pc_corrs(
    pcDF = pcNonCyanoStation, 
    pc = pc, 
    method = "pearson"
  )
}

nonCyanosPearsonCorrected <- ldply(nonCyanosPearson, function(x) x %>% mutate(p = p.adjust(p, method = "fdr")))
nonCyanosPearsonCorrected$Comparison <- rep(c("N1 v N2", "N1 v O", "N2 v O"), 3)

emphasize.strong.cells(which(nonCyanosPearsonCorrected < 0.05, arr.ind = TRUE))
pander(nonCyanosPearsonCorrected, split.tables = Inf)     


```

## Table 2: env variables associated with community dissimilarity
```{r env variables}
library(leaps)

# Function to run a linear model and return the r-squared and pvalue
get_lm <- function(vars, response, dat) {
  formula = reformulate(termlabels = vars, response = response)
  lm.model <- lm(formula, dat)
  r <- summary(lm.model)$adj.r.squared
  p <- summary(lm.model)$coefficients[-1, 4]
  return(c(r,p))
}

# Function to extract liner model adj. r-squared and pvalue
 

get_bestsub_summary <- function(vars, response, dat) {
  formula = reformulate(termlabels = vars, response = response)
  lm.model <- regsubsets(formula, dat)
  bic <- summary(lm.model)$bic
  cp <- summary(lm.model)$cp
  adjr2 <- summary(lm.model)$adjr2
  best.model <- summary(lm.model)$which[which.min(bic), ]
  return(list(model = best.model, bic = bic, cp = cp, adjr2 = adjr2))
}


cyanoEnvVars <- c("Nitrate", "SRP", "Temp", "H2O2", "CDOM", "Ammonia", "Turbidity")

nonCyanoEnvVars <- c("pH", "Turbidity", "Nitrate", "SRP", "DOC", "Temp", "H2O2", "CDOM", "ParMC", "Chla", "Ammonia")

######### Cyanos PC1 ####################
lmSimpleCyanoPC1 <- list()

for (var in cyanoEnvVars) {
  lmSimpleCyanoPC1[[var]] <- get_lm(vars = var, response = "PC1", dat = cyanoDF)
}

########## Cyanos PC2 ######################

lmSimpleCyanoPC2 <- list()

for (var in cyanoEnvVars) {
  lmSimpleCyanoPC2[[var]] <- get_lm(vars = var, response = "PC2", dat = cyanoDF)
}

########## nonCyanos PC1 ##########################
lmSimpleNonCyanoPC1 <- list()

for (var in nonCyanoEnvVars) {
  lmSimpleNonCyanoPC1[[var]] <- get_lm(vars = var, response = "PC1", dat = nonCyanoDF)
}

############ nonCyanos PC2 #####################
lmSimpleNonCyanoPC2 <- list()

for (var in nonCyanoEnvVars) {
  lmSimpleNonCyanoPC2[[var]] <- get_lm(vars = var, response = "PC2", dat = nonCyanoDF)
}

############ nonCyanos PC3 #####################
lmSimpleNonCyanoPC3 <- list()


for (var in nonCyanoEnvVars) {
  lmSimpleNonCyanoPC3[[var]] <- get_lm(vars = var, response = "PC3", dat = nonCyanoDF)
}



rCyano <- rbind(
  "Cyano PC1" = as.data.frame.list(lmSimpleCyanoPC1)[1,], 
  "Cyano PC2" = as.data.frame.list(lmSimpleCyanoPC2)[1,]
)

rNonCyano <- rbind(
  "NonCyano PC1" = as.data.frame.list(lmSimpleNonCyanoPC1)[1,],
  "NonCyano PC2" = as.data.frame.list(lmSimpleNonCyanoPC2)[1,],
  "NonCyano PC3" = as.data.frame.list(lmSimpleNonCyanoPC3)[1,]
)

pCyano <- rbind(
  "Cyano PC1" = as.data.frame.list(lmSimpleCyanoPC1)[2,], 
  "Cyano PC2" = as.data.frame.list(lmSimpleCyanoPC2)[2,]
)

pNonCyano <- rbind(
  "NonCyano PC1" = as.data.frame.list(lmSimpleNonCyanoPC1)[2,],
  "NonCyano PC2" = as.data.frame.list(lmSimpleNonCyanoPC2)[2,],
  "NonCyano PC3" = as.data.frame.list(lmSimpleNonCyanoPC3)[2,]
)
  
pCyano <- t(apply(pCyano, 1, function(x) {p.adjust(x, method = "fdr")}))
emphasize.strong.cells(which(pCyano < 0.05, arr.ind = TRUE))  
pander(rCyano, split.tables = Inf)

write.csv(round(rbind(pCyano, rCyano), digits = 3), "cyano-lm-results.csv")

pNonCyano <- t(apply(pNonCyano, 1, function(x) {p.adjust(x, method = "fdr")}))
emphasize.strong.cells(which(pNonCyano < 0.05, arr.ind = TRUE))  
pander(rNonCyano, split.tables = Inf)

write.csv(round(rbind(pNonCyano, rNonCyano), digits = 3), "noncyano-lm-results.csv")


```


### Best multivariate model for Principal coordinates   
```{r}
# Cyanos
cyPC1 <- get_bestsub_summary(cyanoEnvVars, "PC1", dat = cyanoDF)
cyPC1
cyPC1.fit <- lm(reformulate(cyanoEnvVars[cyPC1$model[-1]], "PC1"), data = cyanoDF)
summary(cyPC1.fit)

cyPC2 <- get_bestsub_summary(cyanoEnvVars, "PC2", dat = cyanoDF)
cyPC2
cyPC2.fit <- lm(reformulate(cyanoEnvVars[cyPC2$model[-1]], "PC2"), data = cyanoDF)
summary(cyPC2.fit)

# NonCyanos
nonCyPC1 <- get_bestsub_summary(nonCyanoEnvVars, "PC1", dat = nonCyanoDF)
nonCyPC1
nonCyPC1.fit <- lm(reformulate(nonCyanoEnvVars[nonCyPC1$model[-1]], "PC1"), data = nonCyanoDF)
summary(nonCyPC1.fit)

nonCyPC2 <- get_bestsub_summary(nonCyanoEnvVars, "PC2", dat = nonCyanoDF)
nonCyPC2
nonCyPC2.fit <- lm(reformulate(nonCyanoEnvVars[nonCyPC2$model[-1]], "PC2"), data = nonCyanoDF)
summary(nonCyPC2.fit)


nonCyanoDFO <- filter(nonCyanoDF, Station == "offshore")

nonCyPC3 <- get_bestsub_summary(nonCyanoEnvVars, "PC3", dat = nonCyanoDFO)
nonCyPC3
nonCyPC3.fit <- lm(reformulate(nonCyanoEnvVars[nonCyPC3$model[-1]], "PC3"), data = nonCyanoDFO)
summary(nonCyPC3.fit)



```

# Figure 6: Alpha diversity
```{r, fig.height=7, fig.width=13}


mytaxa = c("Full", "non-Cyano", "Actinobacteria", "Alphaproteobacteria", "Betaproteobacteria", "Bacteroidetes", "Gammaproteobacteria", "Verrucomicrobia")

############ alpha-proteo #############

act <-
  erieScale %>%
    subset_taxa(Phylum == "Actinobacteria")

alphaprot <-
  erieScale %>%
    subset_taxa(Class == "Alphaproteobacteria")

############ beta-proteo
betaprot <-
  erieScale %>%
    subset_taxa(Class == "Betaproteobacteria")

############ Bacteroidetes ###########
bact <-
  erieScale %>%
    subset_taxa(Phylum == "Bacteroidetes")

gamma <-
  erieScale %>%
    subset_taxa(Class == "Gammaproteobacteria")

verr <-
  erieScale %>%
    subset_taxa(Phylum == "Verrucomicrobia")


full <- erieScale

noncyano <- 
  erieScale %>%
    subset_taxa(Class != "Cyanobacteria")


taxaGroups = list(full, noncyano, act, alphaprot, betaprot, bact, gamma, verr)

obsList <- list()
simpList <- list()

# Loop through each taxonomic group and calculate observed richness and ISI
for (i in 1:8) {
  
  alphaDiv <- plot_richness(
    taxaGroups[[i]], 
    measures = c("Observed", "InvSimpson")
  )

  obsRich <- 
    alphaDiv$data %>% 
    filter(variable == "Observed")

  obsList[[mytaxa[i]]] <- obsRich
  
  invSimp <- 
    alphaDiv$data %>%
    filter(variable == "InvSimpson") 
  
  simpList[[mytaxa[i]]] <- invSimp
  
}

# Function to test whether there is a linear or quadratic relationship 
# between log chla and alpha diversity of a group
test_alphadiv_pp <- function(df, model) {
  
  df <- 
    df %>% 
      select(LogChla, value) %>%
      na.omit()
  
  if (model == "linear") {
    fit <- lm(value ~ LogChla, data = df)
  } else if (model == "quadratic") {
    fit <- lm(value ~ I(LogChla^2), data = df)
  } 
  
  fit.pvalue <- summary(fit)$coef[2,4]
  fit.r2 <- summary(fit)$r.squared
  return(fit.pvalue)
}

# Combine this into one function later
test_alphadiv_r2 <- function(df, model) {
  
  df <- 
    df %>% 
      select(LogChla, value) %>%
      na.omit()
  
  if (model == "linear") {
    fit <- lm(value ~ LogChla, data = df)
  } else if (model == "quadratic") {
    fit <- lm(value ~ I(LogChla^2), data = df)
  } 
  fit.r2 <- summary(fit)$r.squared
  return(fit.r2)
}


linearModelsSimp <- 
  lapply(
    simpList, 
    function(x) test_alphadiv_pp(x, model = "linear")
  ) %>% 
  unlist() %>%
  p.adjust(method = "fdr") %>%
  round(digits = 3)

linearModelsSimpR2 <- 
  lapply(
    simpList, 
    function(x) test_alphadiv_r2(x, model = "linear")
  ) %>% 
  unlist() %>%
  round(digits = 3)

linearModelsObs <- 
  lapply(
    obsList, 
    function(x) test_alphadiv_pp(x, model = "linear")
  ) %>% 
  unlist() %>%
  p.adjust(method = "fdr") %>%
  round(digits = 3)

quadModelsSimp <- 
  lapply(
    simpList, 
    function(x) test_alphadiv_pp(x, model = "quadratic")
  ) %>% 
  unlist() %>%
  p.adjust(method = "fdr") %>%
  round(digits = 3)


# Function to make an alpha-diversity plot with option for fitted line
make_alphadiv_plot <- function(df, measure, group, fit, pvalue, r2) {
  
  g <- 
  ggplot(df, aes(x = LogChla, y = value)) + 
    geom_point() +
    ylab(measure) +
    ggtitle(group) 
  
  if (pvalue != 0) {
    g <- g + annotate(
      "text", 
      x = 1,
      y = max(df$value) - 0.03*max(df$value), 
      size = 3, 
      label = paste("p =", pvalue)
      )
  } else {
    g <- g + annotate(
      "text", 
      x = 1,
      y = max(df$value) - 0.03*max(df$value), 
      size = 3, 
      label = "p < 0.001"
      )
  }
  
  if (pvalue < 0.05) {
    g <- g + annotate(
      "text",
      x = 1,
      y = max(df$value) - 0.08*max(df$value),
      size = 3,
      label = paste("R2 =", r2),
    )
    
  }
  
  
  if (fit == "linear" & pvalue < 0.05) {
    g + geom_smooth(method = "lm", size = 1) 
  } else if (fit == "quadratic") {
    g + geom_smooth(method = "lm", formula = y ~ poly(x, 2), size = 1) 
  } else {
    return(g)
  }
                     
}


simpPlots <- list()

for (i in 1:8) {
  
    simpPlots[[i]] <- make_alphadiv_plot(
      df = simpList[[i]], 
      measure = "InvSimpson", 
      group = mytaxa[i],
      fit = "linear",
      pvalue = linearModelsSimp[i],
      r2 = linearModelsSimpR2[i]
    )

}

obsPlots <- list()

for (i in 1:8) {
  
    obsPlots[[i]] <- make_alphadiv_plot(
      df = obsList[[i]], 
      measure = "Richness", 
      group = mytaxa[i],
      fit = "linear",
      pvalue = linearModelsObs[i],
      r2 = ""
    )

}

grid.arrange(
   obsPlots[[2]], obsPlots[[4]],  obsPlots[[5]], obsPlots[[6]], 
   simpPlots[[2]], simpPlots[[4]],  simpPlots[[5]], simpPlots[[6]],
   ncol = 4
)


grid.arrange(
   obsPlots[[1]], obsPlots[[3]],  obsPlots[[7]], obsPlots[[8]], 
   simpPlots[[1]], simpPlots[[3]],  simpPlots[[7]], simpPlots[[8]],
   ncol = 4
)


```

# Microcystis associates
```{r}

test_htroph_corr <- function(htroph.df, cyano.df, physeq.scale){
  corr <- cor.test( 
                   htroph.df,
                   cyano.df,
                   method = "spearman"
                   )
  
  pvalue <- corr$p
  r <- corr$r
  
  sig <- which(pvalue < 0.05)
  otu.names <- colnames(pvalue)[sig]
  
  sig.tax <- data.frame(tax_table(physeq.scale))[otu.names,]
  sig.p <- pvalue[sig]
  sig.r <- r[sig]
  sig.dat <- data.frame(sig.p, sig.r, sig.tax)
  names(sig.dat)[1:2] <- c("pvalue", "r")
  sig.dat <- arrange(sig.dat, r)
  sig.dat$Genus <- droplevels(sig.dat$Genus)
  return(sig.dat)
}

erie.100 <- subset_samples(erie.subset, Fraction == "100LTR")

# Scale reads to even depth 
erieScale100 <-
  erie.100 %>%
    scale_reads(n = 15000, round = "round") 


n = 15000
thresh = 0.0001

# Prune low abundance taxa using thresh as mean relative abundance
tax_mean <- taxa_sums(erieScale100)/nsamples(erieScale100)
erieScale.0001 <- prune_taxa(tax_mean > thresh*n, erieScale100)

#
nonCyanoN1 <- 
  erieScale.0001 %>%
    subset_taxa(Phylum != "Cyanobacteria") %>%
    subset_samples(Station == "nearshore1") %>% 
    otu_table(.) %>%
    t() %>%
    data.frame()
    
nonCyanoN2 <- 
  erieScale.0001 %>%
    subset_taxa(Phylum != "Cyanobacteria") %>%
    subset_samples(Station == "nearshore2") %>%
    otu_table(.) %>%
    t() %>%
    data.frame()

nonCyanoO <- 
  erieScale.0001 %>%
    subset_taxa(Phylum != "Cyanobacteria") %>%
    subset_samples(Station == "offshore") %>%
    otu_table(.) %>%
    t() %>%
    data.frame()



mcN1 <- 
  erieScale.0001 %>%
    subset_taxa(Species == "Otu00005") %>%
    subset_samples(Station == "nearshore1") %>%
    otu_table(.) %>%
    t() %>%
    data.frame()

 mcN2 <- 
  erieScale.0001 %>%
    subset_taxa(Species == "Otu00005") %>%
    subset_samples(Station == "nearshore2") %>%
    otu_table(.) %>%
    t() %>%
    data.frame()
    
mcO <- 
  erieScale.0001 %>%
    subset_taxa(Species == "Otu00005") %>%
    subset_samples(Station == "offshore") %>%
    otu_table(.) %>%
    t() %>%
    data.frame()



# Run spearman correlation test with full community
corrsPN1 <- list()
corrsRN1 <- list()

for (i in 1:ncol(nonCyanoN1)) {
  corrsPN1[[i]] <- cor.test(nonCyanoN1[,i], 
                            mcN1$Otu00005, method = "spearman")$p.value
  corrsRN1[[i]] <- cor.test(nonCyanoN1[,i], 
                            mcN1$Otu00005, method = "spearman")$estimate
}

#
corrsPN2 <- list()
corrsRN2 <- list()

for (i in 1:ncol(nonCyanoN2)) {
  corrsPN2[[i]] <- cor.test(nonCyanoN2[,i], 
                            mcN2$Otu00005, method = "spearman")$p.value
  corrsRN2[[i]] <- cor.test(nonCyanoN2[,i], 
                            mcN2$Otu00005, method = "spearman")$estimate
}

corrsPO <- list()
corrsRO <- list()

for (i in 1:ncol(nonCyanoO)) {
  corrsPO[[i]] <- cor.test(nonCyanoO[,i], 
                            mcO$Otu00005, method = "spearman")$p.value
  corrsRO[[i]] <- cor.test(nonCyanoO[,i], 
                            mcO$Otu00005, method = "spearman")$estimate
}



corrsCorrectedN1 <-
  corrsPN1 %>%
    unlist() %>%
    p.adjust(method = "fdr")

corrsCorrectedN2 <-
  corrsPN2 %>%
    unlist() %>%
    p.adjust(method = "fdr")

corrsCorrectedO <-
  corrsPO %>%
    unlist() %>%
    p.adjust(method = "fdr")


w.N1 <- which(corrsCorrectedN1 < 0.05)
w.N2 <- which(corrsCorrectedN2 < 0.05)
w.O <- which(corrsCorrectedO < 0.05)

w.int <- intersect(intersect(w.N1, w.N2), w.O)


associates <- data.frame(
  OTU = names(nonCyano.0001)[w.int], 
  RN1 = unlist(corrsRN1)[w.int],
  RN2 = unlist(corrsRN2)[w.int],
  RO = unlist(corrsRO)[w.int]
)


associatesMerge <- left_join(
  associates, 
  data.frame(tax_table(erieScale.0001)), 
  by = c("OTU" = "Species")
)
  




```

# Cyano associates
```{r}
n = 15000
thresh = 0.0001

# Prune low abundance taxa using thresh as mean relative abundance
tax_mean <- taxa_sums(erieScale)/nsamples(erieScale)
eriePrune.0001 <- prune_taxa(tax_mean > thresh*n, erieScale)

cyanoOtuPrune <-
  eriePrune.0001 %>%
  transform_sample_counts(function(x) {x/sum(x)}) %>%
  subset_taxa(Class == "Cyanobacteria") 

library(psych)

cyanoCorrsPearson <- corr.test(t(otu_table(cyanoOtuPrune)), use = "pairwise", method = "pearson", adjust = "fdr")
which(cyanoCorrsPearson$p < 0.05, arr.ind = TRUE)

cyanoCorrsSpearman <- corr.test(t(otu_table(cyanoOtuPrune)), use = "complete", method = "spearman", adjust = "fdr")
which(cyanoCorrsSpearman$p < 0.05, arr.ind = TRUE)

write.csv(cyanoCorrsSpearman$r, "cyano-spearman-r.csv")
write.csv(cyanoCorrsSpearman$p, "cyano-spearman-p.csv")

# Lagged test with syne and microcystis
cyanoOtusTable <- t(otu_table(cyanoOtuPrune))

cor.test(cyanoOtusTable[-(51:53), 1], cyanoOtusTable[-(1:3), 4], method = "pearson")
cco```
